### 2021-03-13 chapters/sections/8-identity_management.tex
überarbeitungen
Anfang und ende fertig

This chapter describes issues of user based identity management in general and for the use case of the OZG. It also presents possibilities of how IMP can be utilized to solve mentioned issues.

\section{OZG Identity Management Issues}

@@

\paragraph{Usability}
Each service provider has a different platform for management of personal information and identity related interactions. This can for example be web pages or applications for desktop and smartphone. Keeping track of all created user profiles is difficult.

@@

To manage personal information stored in each user profile and for example change an e-mail address, users have to access the individual platforms of each service provider. Depending on the amount of created profiles, this can be a time consuming task. Especially if the user has to login to web pages using various authentication methods like passwords, single-sign-on, 2FA or Magic-Link. Users receiving notifications, messages and mails on multiple platforms can quickly loose track and oversee important notifications.

@@

Bidirectional identity related interactions like subscribing for a service, canceling a subscription, sending or receiving a security relevant mail often only takes place trough a web page while e-mails are only used to notify the user to check his user profile ("You have unread messages!"). This is because e-mails do not provide a secure way of exchanging information and don't enable advanced interaction capabilities or exchange of structured content. Although service providers could develop mobile applications as an alternative for the web pages and benefit from secure communication and advanced interaction capabilities, due to the required investments many smaller SPs decide not to.

Service providers rely on the validity of information stored in the user profile. To guarantee that the information is correct, some SPs require users to take extra identification steps by for example performing a video call or requesting the user to show his ID at a local shop. As each service provider manages separate user profiles, the user might have to go trough this time consuming process repeatedly.

Users are not able to automatically share personal information from one user profile with another (except for some Single-Sign-On cases). As a result of that, creating new user profile each time involves entering personal information all over again.

In case of the OZG, usability of user profiles across multiple administration portals is limited: Although user profiles are interoperable, administration portals are individually responsible for the application process. Applying for different administration services, users might have to visit different administration portals. Each of the portals managing the application process separately. Active applications are not stored as part of the user profile but remain accessible only trough the respective portals they were created through. There does not exist an overview of all active applications amongst all administration portals. The user has to go trough each portal to find the application he for example wants to cancel. Especially due to the „one for all“ approach, users will have to access dedicated administration portals for certain types of administrative services. To mitigate this issue, the inbox of user profiles is used to notify the user of the status of all applications.

\paragraph{Data Protection}
In the past, service providers established a habit of collecting more personal information than necessary for providing their services. As a response, the "Datenschutz Grundverordnung" - DSGVO (General Data Protection Regulation) was passed. Amongst other things, it requires service providers to inform the user about which personal information is processed, what the reason for processing the information is and how it is being processed. Conventional user profile systems which store all personal information but do not differentiate between possibly different processing legitimisation of attributes increase the risk of illegitimate data usage.

As a result of the multitude of user profiles and the variety of platforms to manage them, users can quickly loose track of which service provider processes which personal information. Multiple service providers storing sensible information increases the risk of personal information being stolen due to a data breach.

In case of the OZG, the user shares personal information not only with institutions responsible for processing an application but also with multiple systems operated by the member state. Especially the administration portal used to initiate the application is problematic as it receives personal information from the interoperable user profile, sends it to a form server, eventually receives the filled in form and stores it. As a result of that, the administration portal has access to all personal information only the institution will eventually need to process. In addition, institutions actually do not require user profiles for processing applications. After an application is finished, the institution can delete all data stored as part of this process. The only reason user profiles are required is to verify the validity and authorization of incoming applications and to enable the user to comfortably fill in application forms and communicate through an inbox.

@@

User profiles are often secured through passwords. Especially in case of a large amount of profiles, users are tempted to use identical passwords, increasing the risk of it being stolen and all user profiles being accessible. With the amount of different service providers storing personal information and passwords, the risk of it being stolen increases as security of system architectures differs between service providers.

Users also have to rely on the correct operation of the system architecture of the service provider and might not be able to proof failures in the operation of the system. If for example the user cancels a subscription but the system architecture quietly fails to process it, the user might have issues proofing to the service provider, that he did cancel the subscription.

\section{IMP Solution Opportunities}

Identity management provisioning is capable of solving many of the issues described in the previous section.

@@

The goal of IMP solutions is to replace user profiles with a single IMP identity. Through the IMP client a user can manage his IMP identity and all integrated service providers.

Instead of creating user profiles, the user can establish a relationship between his IMP identity and the service provider. Depending on the use case, the relationship can be utilized differently.

Using the IMP client, the user can keep track of all relationships without the need of remembering URLs, usernames or passwords.

It also enables the user to update personal information shared with service providers trough one application.

As the IMP client also contains an inbox, all notifications sent by each service provider can be viewed in the same place.

As part of established relationships, IMP client and service provider can securely exchange bidirectional structured data which along with advanced interaction capabilities of the IMP client can be used to create custom interaction scenarios: Service providers can send structured data to the IMP client along with the information that it should be displayed as a form. Depending on the content of the form and the use case of the service provider, the form can either be a purchase agreement or a survey.

Assumed, through a secure identification and validation process, the identity management provider makes sure, attributes users store as part of their IMP identity are correct, service providers wont require any additional identification steps trough for example video calls.

@@

IMP puts the user back in control of his personal information. As part of the process of establishing a relationship, the service provider has to specify exactly which attributes have to be shared, give a detailed explanation of why the attributes are required to be shared, what the purpose of the relationship is and how the attributes are processed. With this information, the user can make an informed decision whether to establish the relationship and the service provider performed his duty to inform the user as required by the DSGVO.

As personal information is only shared as part of a relationship and with a specified purpose, the service provider has an effective tool to differentiate processing legitimisation of attributes based on the type of relationship they are accessible trough. This helps the service provider in processing personal information according to the DSGVO.

As every relationship along with shared attributes is manageable trough the IMP client, the user wont loose track of which service provider has access to which personal information.

The risk of data being stolen decreases as most of the identity management and data storage is performed by the IMP system and the service provider processes only a small portion of attributes of IMP identities.

In case of the OZG system architecture, IMP can be used by the individual institutions to establish relationships with IMP identities of users to retrieve applications for administrative services. This could eventually eliminate the need of user profiles. Personal information could be shared with only the responsible institution. This solution is described in detail in chapter 5.

@@

The IMP client enables users to access their IMP identity and all relationships without the need of remembering a variety of credentials. This removes the risk of passwords being stolen. The IMP client itself can be secured without the need for a password. As the IMP system is capable of signing relationships and exchanged data, interactions between IMP identity and service provider can be validated at any time. This enables both user and service provider to provide proof for interactions.

>>>>

This chapter describes issues of user based identity management in general and for the specific use case of the OZG. It also presents possibilities of how IMP can be utilized to solve mentioned issues.

\section{Identity Management Issues} \label{section:identity_management_issues}

@@

\paragraph{Usability}
Each Service Provider has a different platform for the management of personal information and identity related interactions. This can for example be web pages or applications for desktop and smartphone. Keeping track of all created user profiles is difficult.

@@

To manage personal information stored in each user profile and for example change an e-mail address, users have to access the individual platforms of each Service Provider. Depending on the amount of created profiles, this can be a time consuming task. Especially if the user has to login to web pages using various authentication methods like passwords, single-sign-on, 2FA or Magic-Link. Users receiving notifications, messages and mails on multiple platforms can quickly loose track and oversee important notifications.

@@

Bidirectional identity related interactions like subscribing to a service, canceling a subscription, sending or receiving a security relevant mail often only takes place through a web page where e-mails are only used to notify the user to check his user profile ("You have unread messages!"). This is because e-mails do not provide a secure way of exchanging information and do not enable advanced interaction capabilities or exchange of structured content. Although Service Providers could develop mobile applications as an alternative for the web pages and benefit from secure communication and advanced interaction capabilities, due to the required investments many smaller SPs decide not to.

Service Providers rely on the validity of information stored in the user profile. To guarantee that the information is correct, some SPs require users to take extra identification steps by for example performing a video call or requesting the user to show their ID at a local shop. As each Service Provider manages separate user profiles, the user might have to go through this time consuming process repeatedly.

Users are not able to automatically share personal information from one user profile to another (except for some Single-Sign-On cases). As a result of that, creating a new user profile each time involves entering personal information all over again.

In case of the OZG, usability of user profiles across multiple administration portals is limited: Although user profiles are interoperable, administration portals are individually responsible for the application process. Applying for different administration services, users might have to visit separate administration portals. Each of the portals manage the application process differently. Active applications are not stored as part of the user profile but remain accessible only through the respective portals they were created through.\\ An overview of all active applications amongst all administration portals does not exist. The user has to go through each portal to find the application they for example want to cancel. Especially due to the „one for all“ approach, users will have to access dedicated administration portals for certain types of administrative services. To mitigate this issue, the inbox of user profiles is used to notify the user of the status of all applications.

\paragraph{Data Protection}
Before, Service Providers established a habit of collecting more personal information than necessary to provide their services. As a response, the "Datenschutz Grundverordnung" - DSGVO (General Data Protection Regulation) was passed. Amongst other things, it requires Service Providers to inform the user about which personal information is processed, the reason for processing the information and how it is being processed. Conventional user profile systems which store all personal information but do not differentiate between possibly different processing legitimisation of attributes increase the risk of illegitimate data usage.

As a result of the multitude of user profiles and the variety of platforms to manage them, users can quickly loose track of which Service Provider processes which personal information. Multiple Service Providers storing sensible information increases the risk of personal information being stolen due to a data breach.

In case of the OZG, the user shares personal information not only with institutions responsible for processing an application but also with multiple systems operated by the member state. Especially the administration portal used to initiate the application is problematic as it receives personal information from the interoperable user profile, sends it to a form server, eventually receives the filled in form and stores it. As a result of that, the administration portal has access to all personal information that in reality only the institution will eventually need to process. In addition, institutions actually do not require user profiles for processing applications. After an application is finished, the institution can delete all data stored as part of this process. The only reason user profiles are required is to verify the validity and authorization of incoming applications and to enable the user to comfortably fill in application forms and communicate through an inbox.

@@

User profiles are most commonly secured through passwords. Especially in case of a large amount of profiles, users are tempted to use identical passwords, increasing the risk of them being stolen and all user profiles being accessible. With the amount of different Service Providers storing personal information and passwords, the risk of it being stolen increases as security of system architectures differs between Service Providers.

Users also have to rely on the correct operation of the system architecture of the Service Provider and might not be able to prove failures in the operation of the system. If for example the user cancels a subscription but the system architecture quietly fails to process it, the user might have issues proving to the Service Provider, that they did cancel the subscription.

\section{IMP Solution Opportunities} \label{section:imp_solution_opportunities}

Identity Management provisioning is capable of solving many of the issues described in the previous section.

@@

The goal of IMP solutions is to replace user profiles with a single IMP identity. Through the IMP client a user can manage their IMP identity and all connected integrated Service Providers. Instead of creating user profiles, the user can establish a relationship between their IMP identity and the Service Provider. Depending on the use case, the relationship can be utilized differently. Using the IMP client, the user can keep track of all relationships without the need of remembering URLs, usernames or passwords. It also enables the user to update personal information shared with Service Providers through one application. As the IMP client also contains an inbox, all notifications sent by each Service Provider can be accessed in the same place.

As part of established relationships, IMP client and Service Provider can securely exchange bidirectional structured data which along with advanced interaction capabilities of the IMP client can be used to create custom interaction scenarios: Service Providers can send structured data to the IMP client along with the information that it should display as a form. Depending on the content of the form and the use case of the Service Provider, the form can for example be a purchase agreement or a survey.

Assuming that, through a secure identification and validation process, the identity management provider has made sure that attributes users store as part of their IMP identity are correct, Service Providers wont require any additional identification steps through for example video calls.

@@

IMP puts the users back in control of their personal information. \\
As part of the process of establishing a relationship, the Service Provider has to specify exactly which attributes have to be shared, give a detailed explanation of why the attributes are required to be shared, what the purpose of the relationship is and how the attributes are processed. With this information, the user can make an informed decision whether to establish the relationship and the Service Provider performed their duty to inform the user as required by the DSGVO.

As personal information is only shared as part of a relationship and with a specified purpose, the Service Provider has an effective tool to differentiate processing legitimisation of attributes based on the type of relationship they are accessible through. This helps the Service Provider in processing personal information according to the DSGVO.

As every relationship along with shared attributes is manageable through the IMP client, the user wont loose track of which Service Provider has access to what personal information.

The risk of data being stolen decreases as most of the identity management and data storage is performed by the IMP system and the Service Provider processes only a small portion of attributes of IMP identities.

In case of the OZG system architecture, IMP can be used by the individual institutions to establish relationships with IMP identities of users to retrieve applications for administrative services. This could eventually eliminate the need for user profiles. Personal information could be shared with only the responsible institution. This solution is described in detail in chapter 5.

@@

The IMP client enables users to access their IMP identity and all relationships without the need of remembering a variety of credentials. This removes the risk of credentials being stolen. The IMP client itself can be secured without the need for a password. As the IMP system is capable of signing relationships and exchanged data, interactions between IMP identity and Service Provider can be validated at any time. This enables both user and Service Provider to provide proof for every interactions.

\section{Outcome}

### 2021-03-13 chapters/sections/9-imp_utilization_in_ozg.tex
Anfang und ende fertig

This chapter presents an IMP solution which solves issues of OZG identity management described in chapter 3. For utilization of IMP services as described by the IMP solution, an integration architecture is presented which enables the OZG system architecture described in section 2.3 to access the IMP system described in section 2.2.

@@

The integration takes place on three layers:

\begin{center}
    \includegraphics[scale=0.6]{Diagrams/Integration Architecture 1/IMP Layer Diagram Integration.pdf}
\end{center}

\paragraph{IMP Solution Layer} 
The IMP solution is integration on the use case level. It presents possibilities of leveraging interaction capabilities of the IMP system in the context of the service provider for improvement of usability, data protection and security. It describes which types of relationships the service provider can use in his business scenario, how relationships can be processed, which types of messages can be exchanged trough the relationships, what the purpose of each message can be and how the IMP client can react to different message types.

@@

\paragraph{Technological Integration} 
The technological integration is integration on the system level. It consists of an integration architecture enabling the existing systems of the service provider to utilize the IMP system as described by the IMP solution. It is separated into a "Communication Integration Layer" and a "Relationship to Identity Integration Layer". On the "Relationship to Identity Integration Layer" the integration architecture enables the existing systems to create relationship templates and establish relationships. On the "Communication Integration Layer" it enables the existing systems to send and receive different types of IMP messages.

Focus of the technological integration is to connect with as few existing systems as possible while solving as many integration issues described in section 3.3 as possible.

>>>>

This chapter presents an IMP solution which solves issues that OZG identity management has, as described in chapter 3. For utilization of IMP services as described by the IMP solution, an integration architecture will be presented that enables the OZG system architecture described in section 2.3 to access the IMP system described in section 2.2.

@@

\begin{figure}[h]
    \centering
    \includegraphics[scale=0.6]{Diagrams/Integration Architecture 1/IMP Layer Diagram Integration.pdf}
    \caption{IMP Layer Diagram}
    \label{integration1:layer_diagram}
\end{figure}

As shown in figure \ref{integration1:layer_diagram}, the integration takes place on three layers:

\paragraph{IMP Solution Layer} 
The IMP solution is integration on the use case level. It presents possibilities of leveraging interaction capabilities of the IMP system in the context of the Service Provider for improvement of usability, data protection and security. Furthermore it describes which types of relationships the Service Provider can use in their business scenario, how relationships can be processed, which types of messages can be exchanged through the relationships, what the purpose of each message can be and how the IMP client can react to different message types.

@@

\paragraph{Technological Integration} 
The Technological Integration is integration on the system level. It consists of an integration architecture enabling the existing systems of the Service Provider to utilize the IMP system as described by the IMP solution. The Technological Integration is separated into a "Communication Integration Layer" and a "Relationship to Identity Integration Layer". On the "Relationship to Identity Integration Layer" the integration architecture enables the existing systems to create relationship templates and establish relationships. On the "Communication Integration Layer" it enables the existing systems to send and receive different types of IMP messages.

Focus of the Technological Integration is to connect with the minimal number systems possibly while solving as many integration issues described in section 3.3 as possible.

### 2021-03-13 msc-thesis.tex
textualizing
message chapter started
messaging chapter done
überarbeitung
überarbeitungen
Anfang und ende fertig

          \vspace{0.5\baselineskip}
          {
               \huge
               \textbf{Integrating Identity Management Providers based on Online Zugangs Gesetz}
          }

@@

@@

@@

\chapter{Introduction}

\section{Context}

\subfile{sections/1-context}

\section{Objective}

\subfile{sections/2-objective}

\section{Structure of Work}

\subfile{sections/3-structure_of_work}

\chapter{Background and Related Work}

\section{Terminology}

\subfile{sections/4-terminology}

\section{Identity Management Provisioning (IMP)}

\subfile{sections/5-imp}

\section{Online Access Law (OZG)}

\subfile{sections/6-online_access_law}

\section{Messaging}

\subfile{sections/7-messaging}

\chapter{OZG Identity Management}

\subfile{sections/8-identity_management}

\chapter{IMP Utilization in OZG}

\subfile{sections/9-imp_utilization_in_ozg}

\chapter{Advanced Utilization}

\subfile{sections/12-advanced_utilization}

\chapter{Solution Evaluation}

\section{Conclusion}

\chapter{Outlook: Advanced IMP Integration}

>>>>

          \vspace{0.5\baselineskip}
          {
               \huge
               \textbf{Integrating Identity Management Providers based on Online Access Law}
          }

@@

Diese Bachelorarbeit stellt Identity Management Provisioning (IMP) als mögliche Lösung von Problemen Benutzerprofil basierter Identitätsmanagementsysteme vor. IMP bietet Identitätsmanagement als Service für Benutzer und Dienstanbieter. Es ermöglicht Benutzern eine IMP-Identität zu erstellen und diese zu nutzen, um persönliche Informationen und Nachrichten mit mehreren Dienstanbietern auszutauschen.

Damit Dienstanbieter mit bestehenden, auf Benutzerprofilen basierenden, Identitätsmanagementsystemen, IMP nutzen können, ist eine Integration in bestehende Geschäftsprozesse und bestehende Systemarchitekturen erforderlich. Es werden zwei IMP-Lösungen vorgestellt, die beschreiben, wie Service-Provider IMP verwenden können. Außerdem wird ein Messaging-System für die technologische Integration präsentiert.

Ziel der IMP-Lösungen ist es, Möglichkeiten zu beschreiben, die Benutzerfreundlichkeit, Datenschutz und Sicherheit des Identitätsmanagements durch die Nutzung des IMP-Dienstes erhöhen. Basierend auf dem aktuell relevanten Anwendungsfall des Online-Zugangs-Gesetzes (OZG) werden zwei IMP-Lösungen vorgestellt. Die erste IMP-Lösung beschreibt eine Möglichkeit, welche Benutzern die Erstellung von Benutzerprofilen, Pflege persönlicher Informationen und Interaktion über eine IMP-Anwendung ermöglicht. Dieser Integrationsansatz soll die Benutzerfreundlichkeit erhöhen und gleichzeitig das Risiko, die Komplexität und die Kosten der Integration minimieren. Dadurch, dass das auf Benutzerprofilen basierende Identitätsmanagement in Betrieb gelassen wird, bleiben Probleme hinsichtlich des Datenschutzes bestehen. Eine zweite IMP-Lösung wird vorgestellt, die den Datenschutz erhöht, indem Benutzerprofile ersetzt werden. Stattdessen geben Benutzer persönliche Informationen für einzelne Geschäftsprozesse temporär frei.

Zur Integration von IMP-Lösungen wird ein Messaging-System vorgestellt, das in der Lage ist, IMP in bestehende Systemarchitekturen von Dienstanbietern zu integrieren. Neben messaging ermöglicht ein modularer Ansatz, dass die Integrationsarchitektur konfigurierbar und erweiterbar ist, um verschiedene IMP-Lösungen in unterschiedliche Systemarchitekturen zu integrieren. Die Fähigkeiten des Messaging-Systems werden durch die Integration der zuvor beschriebenen, grundsätzlich unterschiedlichen, IMP-Lösungen in verschiedene Systemarchitekturen im Kontext des OZG demonstriert.

@@

This bachelor thesis presents Identity Management Provisioning (IMP) as possibility for solving problems of user profile based identity management systems. IMP provides identity management as a service to users and Service Providers. It enables users to create an IMP identity and use it to share personal information and exchange messages with multiple Service Providers.

In order for Service Providers with existing user profile based identity management systems to utilize IMP, integration into their existing business processes and existing system architectures is necessary. Two IMP solutions which describe how Service Providers can utilize IMP and one messaging system for technological integration are presented.

Purpose of IMP solutions is to describe possibilities to increase usability, data protection and security of identity management trough utilization of the IMP service. Based on the currently relevant use case of the Online Access Law (Online Zugangs Gesetz - OZG), two IMP solutions are presented.

The first IMP solution describes the possibility of enabling users to create user profiles, maintain personal information and interact trough an IMP application. This integration approach is designed to increase usability while minimizing risk, complexity and cost of integration.

As a result of leaving user profile based identity management in operation, problems regarding data protection remain. A second IMP solution is presented to increase data protection by replacing user profiles. Instead users temporarily share personal information for individual business processes.

For integration of IMP solutions, a messaging system is presented which is capable of integrating IMP into existing system architectures of Service Providers. In addition to messaging, a modular approach enables the integration architecture to be configurable and expandable to integrate various IMP solutions into different system architectures.

The capabilities of the messaging system are demonstrated by integrating the previously described, fundamentally different, IMP solutions into different system architectures in the context of the OZG.

@@

\chapter{Introduction}
\subfile{chapters/1-introduction}

\chapter{Background and Related Work}
\subfile{chapters/2-background_and_related_work}

\chapter{User Profile Identity Management} \label{chapter:user_profile_identity_management}
\subfile{chapters/3-ozg_identity_management}

\chapter{IMP Utilization as Extension}
\subfile{chapters/4-imp_utilization_in_ozg}

\chapter{IMP Utilization as Replacement}
\subfile{chapters/5-advanced_utilization}

\chapter{Conclusion}
\subfile{chapters/6-conclusion}
